### 2013-02-17 article/paper.tex

\section{Introduction}
The best paper in the world is~\cite{Liben-Nowell:2003:LPP:956863.956972}.

\input link-prediction

\section{Implementation}
The selected algorithms were implemented and tested over three different 
database systems: Relational, Key-Value store and Graph. 
In each system we attempted to implement the algorithms as natively as possible, 
using that system's features and applying common optimizations.

For the bechmarking we chose MySQL, Redis and Neo4J.
These databases are all open-source, enabling us to publish these benchmarks,
and popular in their domain, therefore considered as a viable option for such social network implementation.

\subsection{MySQL}
MySQL is a popular multi-platform, open-source relational database management
system, sponsored by Oracle. It can be used in client-server or embedded
architecture and it's main features include: 
\begin{itemize}
	\item {\bf Pluggable storage-engines:} MySQL supports several storage engines, 
		most notables are MyISM (fast, simple, no ACID support, no foreign keys) and
		InnoDB (default, supports transactions, ACID and foreign keys).
	\item {\bf Stored Procedures} Like most other relational databases MySQL supports 
		stored procedures execution. Advantages of this feature include 
		avoidance of network traffic for the SQL queries and intermediate results,
		and encapsulation of a multi-query logic in one place.
	\item {\bf Fast Bulk Data Loading} MySQL support very fast data loading from file,
		with it's 'LOAD DATA' statement. From all the systems we tested, this offered 
		the easiest and fastest to load our graph.
\end{itemize}

In this case, the algorithms were implemented as a SQL stored-procedures only,
using indices and intermediate tables when possible. A Python script was used
to run these stored-procedures and collect the benchmarks. For a storage-engine 
we chose InnoDB, the default engine in MySQL 5.5, as it out-performs MyISM in 
most senarios. \footnote{Sara: can we cite Oracle white papers? http://www.oracle.com/partners/en/knowledge-zone/mysql-5-5-innodb-myisam-522945.pdf}

\subsection{Redis}
Redis is an open-source, in-memory key-value store database system, sponsored
by VMware. It's main features are atomic operations, data structure as value
(hash, list, set and sorted-set), optional persistence, pub/sub mechanism,
built-in Lua interpreter and master-slave replication.

Here, some of the algorithms were implemented in Python using redis API, and
some were implemented as Lua scripts (to minimize Redis API calls). When an
index was needed, we simply used another Redis DB (you get 16 with the default
configuration). 
The graphs were represented using adjacency lists, i.e. a key is a node-id and
it's value is a set of his neighbors. That implementation posed a limitation,
since we couldn't store nodes without neighbors, as redis doesn't allow to
store empty sets. \#TODO: why doesn't it matter in our case

\subsection{Neo4J}
Neo4J is a multi-platform, open-source, graph database supported by Neo
Technology. It features full ACID transactions, expressive graph query
language, indexing and high availability. It can operate in an embedded or
client-server architecture.

In this database, we attempted to use only Neo4j's declarative query language
(Cypher) to implement the algorithms, and the client-server architecture to run
them with. The benchmarking was performed using the Python API, which also
assisted in implementing some multi-query algorithms. We found that the best
way to load the graph into Neo4J was to use the BatchInserter class in the
embedded architecture, and copy the created files into the server.

There are some limitations to the Cypher language users should be aware of:
\begin{itemize}
\item There are not many built in functions (e.g. no Power, Log, or Rand
  functions) 
\item No support for user defined functions
\item Queries are really slow when matching a non-trivial pattern over the whole
graph (e.g. common neighbors)  
\end{itemize}

\section{Experiments}
We benchmarked the above implementations using 10 undirected graphs:
\begin{itemize}
\item Three graphs were extracted from the DBLP XML records using the authors as
nodes, and co-authorship relation as edges. Each graph was over a different
time period, and  we only kept authors with 3 or more publication within the
selected time period (core-3).

\item Seven graphs from the Stanford Network Analysis Project (SNAP) dataset
  collection. No changes were made to these graphs.
\end{itemize}

Graphs to add:
\begin{enumerate}
\item Top N index generation (significantly faster on MySQL)
\item Graph Distance (significantly faster on Redis)
\item Jaccard Coefficient (faster on Neo4J)
\end{enumerate}

\section{Conclusions}
The main conclusion we took from this experience is that there is no silver-bullet for storing and running algorithms over social-network graphs. Generally speaking, MySQL seems to have an edge with indices-based implementation, Redis seems to be extremely flexible and open to endless optimization, and Neo4J's Cypher is leading with implementation simplicity.
If we were to build a social network database we'd consider holding replicas in server types of system. While this solution might not be disk-space efficient, it'll provide us the opportunity to achieve the fastest query time with minimal time investment.

Please note that we didn't use any scale out features. A large network implementation can't pick a solution which doesn't have it. \#TODO